---
title: "EyetrackingDataCleanUp"
author: "RF"
date: "2/1/2018"
output: html_document
---

# Here we load libraries and data (from 2018)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(readr,groupdata2,ggplot2,tidyverse,grid,jpeg)

# Both experiments are in the same files
setwd("C:/Users/Viki/Documents/egyetem/komp.mod/eye-tracking workshop/eye-tracking/new")
Samples=read_delim('eyetracking_2018_samples.txt',delim='\t')
Fixations=read_delim('eyetracking_2018_fixations.txt',delim='\t')
Saccades=read_delim('eyetracking_2018_saccades.txt',delim='\t')
```

# Here we clean up the data
- We only select useful columns
- We rename the columns in more concise and meaningful ways
- We add the task (Visual Search vs. Social Engagement)
- We add time as counted from the start of each trial
- We check whether the trials are of the length we expect (6s for videos, 10s/20s/40s for images)

```{r}

AddTask <- function(df){
  for (s in unique(df$ParticipantID)){
    if (max(df$Trial[df$ParticipantID==s])==8){
      df$Task[df$ParticipantID==s]='SocialEngagement'
    } else {
      df$Task[df$ParticipantID==s]='VisualSearch'
    }
  }
  df$ParticipantGender[grepl("f|F",df$ParticipantID)]='female'
  df$ParticipantGender[grepl("m|M",df$ParticipantID)]='male'
  
  df$SearchOrder[grepl("f1|F1",df$ParticipantID)]=1
  df$SearchOrder[grepl("m1|M1",df$ParticipantID)]=1
  df$SearchOrder[grepl("f2|F2",df$ParticipantID)]=2
  df$SearchOrder[grepl("m2|M2",df$ParticipantID)]=2
  
  return(df)
}

Fixations = Fixations %>% 
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Fixation=CURRENT_FIX_INDEX,
    Duration=CURRENT_FIX_DURATION,
    PupilSize=CURRENT_FIX_PUPIL,
    PositionX=CURRENT_FIX_X,
    PositionY=CURRENT_FIX_Y,
    StartTime=CURRENT_FIX_START,
    EndTime=CURRENT_FIX_END)

Fixations$PositionYClean=1081-Fixations$PositionY

Fixations=AddTask(Fixations)


Saccades = Saccades %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Saccade=CURRENT_SAC_INDEX,
    Direction=CURRENT_SAC_DIRECTION,
    Amplitude=CURRENT_SAC_AMPLITUDE,
    Duration=CURRENT_SAC_DURATION,
    PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
    MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
    Blink=CURRENT_SAC_CONTAINS_BLINK,
    StartTime=CURRENT_SAC_START_TIME,
    EndTime=CURRENT_SAC_END_TIME,
    StartX=CURRENT_SAC_START_X,
    StartY=CURRENT_SAC_START_Y,
    EndX=CURRENT_SAC_END_X,
    EndY=CURRENT_SAC_END_Y
  )
Saccades$Amplitude=as.numeric(Saccades$Amplitude)
Saccades$MeanVelocity=as.numeric(Saccades$MeanVelocity)
Saccades$EndX=as.numeric(Saccades$EndX)
Saccades$EndY=as.numeric(Saccades$EndY)
Saccades$StartYClean=1081-Saccades$StartY
Saccades$EndYClean=1081-Saccades$EndY

Saccades=AddTask(Saccades)

Samples = Samples %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    TimeStamp=TIMESTAMP,
    PositionX=RIGHT_GAZE_X,
    PositionY=RIGHT_GAZE_Y,
    PupilSize=RIGHT_PUPIL_SIZE,
    Saccade=RIGHT_IN_SACCADE,
    Blink=RIGHT_IN_BLINK,
    MeanVelocityX=RIGHT_VELOCITY_X,
    MeanVelocityY=RIGHT_VELOCITY_Y,
    MeanAccellerationX=RIGHT_ACCELLERATION_X,
    MeanAccellerationY=RIGHT_ACCELLERATION_Y
  )

Samples$MeanVelocityY=as.numeric(Samples$MeanVelocityY)
Samples$MeanVelocityX=as.numeric(Samples$MeanVelocityX)
Samples$MeanAccellerationY=as.numeric(Samples$MeanAccellerationY)
Samples$MeanAccellerationX=as.numeric(Samples$MeanAccellerationX)
Samples$PositionYClean=1081-Samples$PositionY

Samples=AddTask(Samples)

# Add time from trial start

Samples <- Samples %>%
  group_by(ParticipantID,Trial) %>%
  mutate(TrialTime = TimeStamp - min(TimeStamp))

## Double check trial timing. It's all in place!
X <- Samples %>%
   group_by(ParticipantID,Trial) %>%
   dplyr::summarize(LengthSeconds = max(TrialTime,na.rm=T)/1000)
print(X)
```

## Import fixation number and saccade number from processed data to samples
Since the samples do not exclude artifacts and do not contain fixation and saccade number, we import this data from Fixations and Saccades datasets

```{r}

for (t in seq(nrow(Fixations))){
  StartTime=Fixations$StartTime[t]
  EndTime=Fixations$EndTime[t]
  ID=Fixations$ParticipantID[t]
  Trial=Fixations$Trial[t]
  FixN=Fixations$Fixation[t]
  Samples$Fixation[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=FixN
}


for (t in seq(nrow(Saccades))){
  StartTime=Saccades$StartTime[t]
  EndTime=Saccades$EndTime[t]
  ID=Saccades$ParticipantID[t]
  Trial=Saccades$Trial[t]
  SacN=Saccades$Fixation[t]
  Samples$SaccadeN[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=SacN
}


setwd("C:/Users/Viki/Documents/egyetem/komp.mod/eye-tracking workshop/eye-tracking/new")

write_csv(Saccades,"SaccadesV1.csv")
write_csv(Fixations,"FixationsV1.csv")
write_csv(Samples,"SamplesV1.csv")

```

## Merge with data about experimental conditions
We merge the data with the experimental logs.

```{r}
setwd("C:/Users/Viki/Documents/egyetem/komp.mod/eye-tracking workshop/eye-tracking/new")

Fixations=read_csv('FixationsV1.csv')
Saccades=read_csv('SaccadesV1.csv')
Samples=read_csv('SamplesV1.csv')

## THEN GO TO THE LOG FILES AND MATCH the data from the logfiles to the samples/fixations/saccades (to have conditions, and so forth)
## ADD GROUP PARTICIPANT AND GENDER TO THE TABLES

Order1<- read_tsv('ForagingOrder.txt',col_names = F)
Order1$Trial=1:10
Order1$SearchOrder=1
Order1$SearchType=c("Search","Search","Search","Search","Search","Count","Count","Count","Count","Count")
Order2<- read_tsv('ForagingOrder.txt',col_names = F)
Order2$Trial=1:10
Order2$SearchOrder=2
Order2$SearchType=c("Count","Count","Count","Count","Count","Search","Search","Search","Search","Search")
Order=rbind(Order1,Order2)
Order=Order %>% rename(Stimulus=X1)

Saccades=merge(Saccades,Order,all=T)
Fixations=merge(Fixations,Order,all=T)
Samples=merge(Samples,Order,all=T)

setwd("C:/Users/Viki/Documents/egyetem/komp.mod/eye-tracking workshop/eye-tracking/new/logfiles")
fileList<-list.files(".",pattern="csv")
logs=NULL
for (f in fileList){
  d<-read_csv(f)
  d=d %>% rename(Trial=X1,ParticipantID=subject)
  d$Trial=d$Trial+1
  for (n in seq(nrow(d))){
    x<-str_split(d$video[n],"_")[[1]]
    d$ActorGender[n]=x[1]
    d$Directionality[n]=x[4]
    d$Ostension[n]=x[5]
  }
  if (exists('logs')){logs=rbind(logs,d)} else {logs=d}
}

Saccades=merge(Saccades,logs,all=T)
Saccades$Stimulus[Saccades$Task=='SocialEngagement']=Saccades$video[Saccades$Task=='SocialEngagement']
Saccades$video=NULL
Fixations=merge(Fixations,logs,all=T)
Fixations=Fixations[!is.na(Fixations$Task),]
Fixations$Stimulus[Fixations$Task=='SocialEngagement']=Fixations$video[Fixations$Task=='SocialEngagement']
Fixations$video=NULL
Samples=merge(Samples,logs,all=T)
Samples$Stimulus[Samples$Task=='SocialEngagement']=Samples$video[Samples$Task=='SocialEngagement']
Samples$video=NULL
setwd("C:/Users/Viki/Documents/egyetem/komp.mod/eye-tracking workshop/eye-tracking/new")
write_csv(Saccades,"SaccadesV2.csv")
write_csv(Fixations,"FixationsV2.csv")
write_csv(Samples,"SamplesV2.csv")
```

# We now integrate the datasets with all the available data from 2017

```{r}

Fixations=read_csv('FixationsV2.csv')
Saccades=read_csv('SaccadesV2.csv')
Samples=read_csv('SamplesV2.csv')

ForagingFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_FixationsClean.csv")
ForagingFixations$Task='VisualSearch'
ForagingSaccades <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_SaccadesClean.csv")
ForagingSaccades$Task='VisualSearch'
PupilsFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_FixationsClean.csv")
PupilsFixations$Task='SocialEngagement'
PupilsSamples <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_SamplesClean.csv")
PupilsSamples$Task='SocialEngagement'


ForagingFixations <- ForagingFixations %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Duration=Duration,
  StartTime=CURRENT_FIX_START,
  EndTime=CURRENT_FIX_END,
  PositionX=CURRENT_FIX_X,
  PositionY=CURRENT_FIX_Y
)

ForagingFixations[,c('PupilSize','ActorGender','Directionality','Ostension','ParticipantGender')]=NA

PupilsFixations <- PupilsFixations %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  Fixation=current_fix_index,
  Duration=current_fix_duration,
  StartTime=current_fix_start,
  EndTime=current_fix_end,
  PositionX=current_fix_x,
  PositionY=current_fix_y,
  Task=Task,
  PupilSize=current_fix_pupil,
  Stimulus=video,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension
)

PupilsFixations[,c('Condition','SearchType','SearchOrder','ParticipantGender')]=NA

Fixations=Fixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

ForagingFixations=ForagingFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

PupilsFixations=PupilsFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]


Fixations2017=rbind(ForagingFixations,PupilsFixations)
Fixations2017$Year='2017'
Fixations$Year='2018'
Fixations=rbind(Fixations,Fixations2017)

ForagingSaccades <- ForagingSaccades %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Direction=CURRENT_SAC_DIRECTION,
  Amplitude=CURRENT_SAC_AMPLITUDE,
  PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
  MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
  Blink=CURRENT_SAC_CONTAINS_BLINK,
  StartTime=CURRENT_SAC_START_TIME,
  EndTime=CURRENT_SAC_END_TIME,
  StartX=CURRENT_SAC_START_X,
  StartY=CURRENT_SAC_START_Y,
  EndX=CURRENT_SAC_END_X,
  EndY=CURRENT_SAC_END_Y
)

ForagingSaccades[,c("ParticipantGender","Stimulus","ActorGender","Directionality","Ostension" )]=NA
ForagingSaccades$Year='2017'
Saccades=Saccades[,c("ParticipantID","Trial","Saccade","Direction","Amplitude","PeakVelocity","MeanVelocity","StartTime","EndTime","Duration", "StartX", "StartY","EndX","EndY","Blink","SearchType","SearchOrder","Task","ParticipantGender","Stimulus", "ActorGender" , "Directionality","Ostension")]
Saccades$Year='2018'  
Saccades=rbind(Saccades,ForagingSaccades)

PupilsSamples <- PupilsSamples %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  PupilSize=right_pupil_size,
  Fixation=fix_nr,
  Saccade=right_in_saccade,
  Blink=right_in_blink,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension,
  Stimulus=video,
  TrialTime=task_time,
  Task=Task
)

PupilsSamples[,c("SearchOrder","TimeStamp","PositionX","PositionY","MeanVelocityX","MeanVelocityY","MeanAccellerationX","MeanAccellerationY","ParticipantGender","SearchType")]=NA
PupilsSamples$fixation=NULL

Samples=Samples[,colnames(PupilsSamples)] 
PupilsSamples$Year='2017'
Samples$Year='2018'
Samples=rbind(Samples,PupilsSamples)

Fixations$Ostension[Fixations$Ostension=='+o']='Ostensive'
Fixations$Ostension[Fixations$Ostension=='-o']='Private'
Samples$Ostension[Samples$Ostension=='+o']='Ostensive'
Samples$Ostension[Samples$Ostension=='-o']='Private'

Fixations$Directionality[Fixations$Directionality=='div']='Towards3rd'
Fixations$Directionality[Fixations$Directionality=='dir']='TowardsYou'
Samples$Directionality[Samples$Directionality=='div']='Towards3rd'
Samples$Directionality[Samples$Directionality=='dir']='TowardsYou'

Fixations$ActorGender[Fixations$ActorGender=='f']='Female'
Fixations$ActorGender[Fixations$ActorGender=='m']='Male'
Samples$ActorGender[Samples$ActorGender=='f']='Female'
Samples$ActorGender[Samples$ActorGender=='m']='Male'

Fixations$SearchOrder[Fixations$SearchOrder=='First']='1'
Fixations$SearchOrder[Fixations$SearchOrder=='Second']='2'
Saccades$SearchOrder[Saccades$SearchOrder=='First']='1'
Saccades$SearchOrder[Saccades$SearchOrder=='Second']='2'

write_csv(Saccades,"SaccadesV3.csv")
write_csv(Fixations,"FixationsV3.csv")
write_csv(Samples,"SamplesV3.csv")

```


## Now plotting the data (only 2018 data here, use V3 if you want to visualize them all)

```{r}
SamplesV2=read_csv('SamplesV2.csv')
FixationsV2=read_csv('FixationsV2.csv')
SaccadesV2=read_csv('SaccadesV2.csv')

#SANITY CHECKS

#check for outliers
ggplot(FixationsV2, aes(ParticipantID,Duration)) +
  geom_boxplot() +
  facet_grid(.~Task)

#check distribution of fixations
hist(FixationsV2$Duration[FixationsV2$Task=='VisualSearch']) #lognormal-ish
hist(FixationsV2$Duration[FixationsV2$Task=='SocialEngagement']) #lognormal-ish

ggplot(FixationsV2, aes(Duration,color=Task)) + geom_density()

#check distribution of saccades
hist(SaccadesV2$Amplitude[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$Amplitude[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish
hist(SaccadesV2$MeanVelocity[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$MeanVelocity[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish
hist(SaccadesV2$PeakVelocity[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$PeakVelocity[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish

#check distribution of pupil size 
hist(SamplesV2$PupilSize[SamplesV2$Task=='SocialEngagement']) #multimodal?
ggplot(subset(SamplesV2,Task=='SocialEngagement'),aes(PupilSize,color=ParticipantID))+geom_density() # Not really!
hist(SamplesV2$PupilSize[SamplesV2$Task=='VisualSearch']) #bimodal?
ggplot(subset(SamplesV2,Task=='VisualSearch'),aes(PupilSize,color=ParticipantID))+geom_density() # Not really!

#RUN SOME ANALYSES AND VISUALIZATIONS

#visualize VELOCITY over trial time for ONE participant/trial
ggplot(data=subset(SamplesV2,ParticipantID=='1_1_f1' & Trial==1), aes(x=TrialTime, y=abs(MeanVelocityX))) +
  geom_line()+
  geom_point() # + facet_grid(.~ParticipantGender)

#visualize ACCELLERATION over trial time for ONE participant/trial
ggplot(data=subset(SamplesV2,ParticipantID=='1_1_f1' & Trial==1), aes(x=TrialTime, y=abs(MeanAccellerationX))) +
  geom_line()+
  geom_point() # + facet_grid(.~ParticipantGender)

#visualize SACCADE AMPLITUDE (difference between VisualSeach and SocialEngagement in terms of saccades)
ggplot(data=SaccadesV2, aes(x=Saccade,y=Amplitude,group=ParticipantID)) +
  geom_line(aes(color=ParticipantID))+
  geom_point(aes(color=ParticipantID))+
  facet_grid(.~Task)

#visualize FIXATION DURATION (difference between VisualSeach and SocialEngagement in terms of fixations)
ggplot(data=FixationsV2, aes(x=Fixation,y=Duration,group=ParticipantID)) +
  geom_line(aes(color=ParticipantID))+
  geom_point(aes(color=ParticipantID))+
  facet_grid(.~Task)

#visualize FIXATION DURATION BY GENDER
ggplot(data=subset(FixationsV2, Trial==1), aes(x=Fixation,y=Duration)) +
  geom_point()+
  geom_line()+
  geom_smooth() +
  facet_grid(.~ParticipantGender)


# HEATMAPS (https://gist.github.com/kalebr/6832380)

jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
img <- readJPEG('eyetrackingscripts/Foraging/ng021ws.jpg')
g <- rasterGrob(img, interpolate=TRUE)

#density
ggplot(subset(FixationsV2, Task=='VisualSearch' & ParticipantID=='6_3_m2' & Trial==1), aes(x = PositionX, y = PositionYClean)) +
  xlim(0,1920) +
  ylim(0, 1080) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=0, ymax=1080) + #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  stat_density2d(geom="raster", aes(fill=..density.., alpha=sqrt(sqrt(..density..))), contour=FALSE, n=1000) + 
  scale_alpha(range = c(0.1, 0.6)) + scale_fill_gradientn(colours = jet.colors(10), trans="sqrt")

x=subset(FixationsV2, ParticipantID=='6_3_m2' & Trial==1)
x= x[order(x$Fixation),]

#SCANPATHS
ggplot(x, aes(x=PositionX, y=PositionYClean, label=Fixation)) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-0, ymax=1080) + #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  geom_point(size = x$Duration[x$ParticipantID=='6_3_m2' & x$Trial==1]/100, alpha = .5, color='red') +
  geom_path(size = 1, alpha = 0.3,color='red') +
  geom_text(aes(label = Fixation, size = 5))

#GROWTH PLOTS PUPIL SIZE
ggplot(subset(SamplesV2, Task=='SocialEngagement'),   #evt. FixationsV2,StartTime
       aes(TrialTime, PupilSize, colour=ParticipantGender), na.rm=T) +
       geom_smooth() +
       xlim(0,5371)+
       facet_grid(.~ActorGender) #Task #Ostension #ActorGender #Directionality


```

